\documentclass{mathnotes}
\usepackage{tikz-cd} % Commutative diagrams

\name{Jacky Lee}
\notetitle{Category Theory Notes}
\notedate{\today}

\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\cod}{cod}
\DeclareMathOperator{\id}{id}
\newcommand{\cat}[1]{\text{\textbf{#1}}}

\begin{document}
\begin{center}
  \vspace*{20pt}
  \LARGE{Category Theory Notes}
\end{center}

\begin{defi}
  A \define{metagraph} consists of \define{objects} $a, b, c, \ldots$,
  \define{arrows} $f, g, h, \ldots$, and two operations as follows:
  \begin{itemize}
    \item \define{Domain}, which assigns to each arrow $f$ an object $a=\dom f$
    \item \define{Codomain}, which assigns to each arrow $f$ an object $b=\cod
      f$
  \end{itemize}
  These operations on $f$ are usually indicated by
  \[f:a\rightarrow b\hspace{10pt}\text{or}\hspace{10pt}a\xrightarrow{f}b\]
\end{defi}

\begin{defi}
  A \define{metacategory} is a metagraph with two additional operations:
  \begin{itemize}
    \item \define{Identity}, which assigns to each object $a$ an arrow
      $\id_a=1_a:a\rightarrow a$
    \item \define{Composition}, which assigns to each pair $\langle g,f
      \rangle$ of arrows with $\dom g=\cod f$ an arrow $g\circ f$, called their
      \define{composite}, with $g\circ f:\dom f\rightarrow\cod g$
  \end{itemize}
  This operation may be pictured by the diagram
  \begin{center}
    \begin{tikzcd}
      A\arrow[r, "f"]\arrow[dr, "g\circ f", swap] & B\arrow[d, "g"] \\
                                                  & C
    \end{tikzcd}
  \end{center}
  These operations in a metacategory are subject to the two following axioms:
  \begin{itemize}
    \item \define{Associativity}. For given objects and arrow in the
      configuration
      \[a\xrightarrow{f}b\xrightarrow{g}c\xrightarrow{k}d\]
      one always has the equality
      \[k\circ\left(g\circ f\right)=\left(k\circ g\right)\circ f\]
      This axiom is represented by the statement that the following diagram is
      commutative.
      \begin{center}
        \begin{tikzcd}[column sep=3cm, row sep=huge]
          A\arrow[r, "k\circ (g\circ f) = (k\circ g)\circ f"]
          \arrow[d, "f"']\arrow[dr, "g\circ f", pos=0.2, swap]
          & D \\
          B\arrow[r, "g", swap]
          \arrow[ur, "k\circ g", pos=0.8, swap, crossing over]
          & C\arrow[u, "k"']
        \end{tikzcd}
      \end{center}
    \item \define{Unit law}. For all arrows $f:a\rightarrow b$ and
      $g:b\rightarrow c$, composition with the identity arrow $1_b$ gives
      \[1_b\circ f=f\hspace{10pt}\text{and}\hspace{10pt}g\circ1_b=g\]
      This axiom is represented by the statement that the following diagram is
      commutative.
      \begin{center}
        \begin{tikzcd}
          A\arrow[r, "f"]\arrow[dr, "f", swap]
                    & B\arrow[d, "1_b"]\arrow[dr, "g"', swap] \\
                    & B\arrow[r, "g"'] & C
        \end{tikzcd}
      \end{center}
  \end{itemize}
\end{defi}

\begin{defi}
  A \define{directed graph} is a set $O$ of objects, a set $A$ of arrows, and
  two functions
  \[A\xrightarrow{\dom}O\hspace{50pt}A\xrightarrow{\cod}O\]
  The set of composable pairs of arrows is the set
  \[A\times_OA=\{\langle g,f\rangle\ |\ g,f\in A
  \text{ and }\dom g=\cod f\}\]
\end{defi}

\begin{bdefi}
  A \define{category} is a directed graph with two additional functions
  \[O\xrightarrow{\id}A,\ A\times_OA\xrightarrow{\circ}A,\]
  \[c\mapsto\id_c,\ \langle g,f\rangle\mapsto g\circ f\]
  called identity and composition (also written as $gf$), such that
  \[\dom(\id a)=a=\cod(\id a),\hspace{20pt}\dom(g\circ
  f)=\dom f,\hspace{20pt}\cod(g\circ f)=\cod g\]
  for all objects $a\in O$ and all composable pairs of arrows $\langle
  g,f\rangle\in A\times_OA$, and such that the associativity and unit axioms
  hold.
\end{bdefi}

\begin{note}
  We usually drop the letters $A$ and $O$, and write
  \[c\in C\hspace{50pt}f\text{ in }C\]
  for ``$c$ is an object of $C$'' and ``$f$ is an arrow of $C$'', respectively.
\end{note}

\begin{defi}
  We write
  \[\hom(b,c)=\{f\ |\ f\text{ in }C,\dom f=b, \cod f=c\}\]
  for the set of arrows from $b$ to $c$.
\end{defi}

\begin{ex}
  The following are simple examples of categories.
  \begin{itemize}
    \item \textbf{0} is the empty category with no objects or arrows.
    \item \textbf{1} is the category with one object and one identity arrow.
    \item \textbf{2} is the category $A\rightarrow B$ with two objects and one
      non-identity arrow.
    \item \textbf{3} is the category with three objects whose non-identity
      arrows are arranged as in the following triangle.
      \begin{center}
        \begin{tikzcd}
          A\arrow[r]\arrow[dr, swap] & B\arrow[d] \\
                                     & C
        \end{tikzcd}
      \end{center}
    \item $\downdownarrows$ is the category $A\rightrightarrows B$ with two
      objects and two non-identity arrows. We call two such arrows
      \define{parallel arrows}.
  \end{itemize}
\end{ex}

\begin{bdefi}
  A \define{functor} is a morphism of categories. For categories $C$ and $B$, a
  functor $T:C\rightarrow B$ consists of two related functions: The
  \define{object function} $T$, which assigns to each object $c$ of $C$ an
  object $Tc$ of $B$ and the \define{arrow function} (also written $T$) which
  assigns to each arrow $f:c\rightarrow c'$ of $C$ an arrow $Tf:Tc\rightarrow
  Tc'$ of $B$, in such a way that
  \[T(1_c)=1_{Tc}\hspace{50pt} T(g\circ f)=Tg\circ Tf\]
\end{bdefi}

\begin{defi}
  The \define{forgetful functor} (or \define{underlying functor}) is a functor
  which ``forgets'' some of all of the structure of an algebraic object.
\end{defi}

\begin{ex}
  The functor $U: \cat{Rng}\rightarrow\cat{Grp}$ that assigns to each ring $R$
  the set $UR$ of its elements and assigns to each morphism $f:R\rightarrow R'$
  of rings the same function $f$, regarded just as a group homomorphism rather
  than a ring homomorphism (thus ``forgetting'' the multiplication).
\end{ex}

\begin{defi}
  The \define{composite} of two functors $T:C\rightarrow B$ and $S:B\rightarrow
  A$ is $S\circ T:C\rightarrow A$.
\end{defi}

\begin{defi}
  An \define{isomorphism} $T:C\rightarrow B$ of categories is a functor $T$
  from $C$ to $B$ which is a bijection, both on objects and on arrows.
  Alternatively, a functor $T:C\rightarrow B$ is an isomorphism if and only if
  there exists a functor $S:B\rightarrow C$ for which both composites $S\circ
  T$ and $T\circ S$ are identity functors. Then, $S$ is the \define{two-sided
  inverse} $S=T^{-1}$.
\end{defi}

\begin{defi}
  A functor $T:C\rightarrow B$ is \define{full} when to every pair $c, c'$ of
  objects of $C$ and to every arrow $g:Tc\rightarrow Tc'$ of $B$, there exists
  an arrow $f:c\rightarrow c'$ of $C$ with $g=Tf$.
\end{defi}

\begin{defi}
  A functor $T:C\rightarrow B$ is \define{faithful} when to every pair $c, c'$
  of objects of $C$ and to every pair $f_1,f_2:c\rightarrow c'$ of parallel
  arrows of $C$, the equality $Tf_1=Tf_2:Tc\rightarrow Tc'$ implies $f_1=f_2$.
\end{defi}

For a pair of objects $c,c'\in C$, the arrow function of $T:C\rightarrow B$
assigns to each $f:c\rightarrow c'$ an arrow $Tf:Tc\rightarrow Tc'$ and so we
can define
\[T_{c,c'}:\hom(c,c')\rightarrow\hom(Tc,Tc'),\hspace{50pt}f\mapsto Tf\]
Then $T$ is full when every such function is surjective, and faithful when
every such function is injective. For a functor which is both full and
faithful, every such function is a bijection, but this need not mean that the
functor itself is an isomorphism of categories, for there may be objects of $B$
not in the image of $T$.

\begin{defi}
  A \define{subcategory} $S$ of a category $C$ is a collection of some of the
  objects and some of the arrows of $C$, which includes with each arrow $f$
  both the object $\dom f$ and the object $\cod f$, with each object $s$ its
  identity arrow $1_s$ and with each pair of composable arrows $s\rightarrow
  s'\rightarrow s''$ their composite.
\end{defi}

\begin{defi}
  Given a subcategory $S$ of $C$, the \define{inclusion functor} is the functor
  $T:S\rightarrow C$ which sends each object and each arrow of $S$ to itself in
  $C$.
\end{defi}

\begin{defi}
  $S$ is a \define{full subcategory} of $C$ when the inclusion functor
  $T:S\rightarrow C$ is full.
\end{defi}

\begin{bdefi}
  Given two functors $S,T:C\rightarrow B$, a \define{natural transformation}
  $\tau:S\xrightarrow{\bullet}T$ is a function which assign to each object $c$
  of $C$ an arrow $\tau_c=\tau c:Sc\rightarrow Tc$ of $B$ in such a way that
  every arrow $f:c\rightarrow c'$ in $C$ yields a diagram
  \begin{center}
    \begin{tikzcd}
      c\arrow[d, "f"] & Sc\arrow[r,"\tau c"]\arrow[d,"Sf"] &
      Tc\arrow[d,"Tf"] \\
      c'              & Sc'\arrow[r,"\tau c'"]             &
      Tc'
    \end{tikzcd}
  \end{center}
  which is commutative. When this holds, we say that $\tau_c:Sc\rightarrow Tc$
  is \define{natural} in $c$. We call $\tau a,\tau b,\tau c,\ldots$ the
  \define{components} of the natural transformation $\tau$.
\end{bdefi}

\begin{defi}
  A natural transformation $\tau$ with every component $\tau c$ invertible in
  $B$ is called a\\
  \define{natural equivalence} or a \define{natural isomorphism}. Symbolically,
  $\tau:S\cong T$.
\end{defi}

\end{document}
